---
title: "DASC 5420 - Assignment 1"
author: "Pham Thi Thai - T00727094"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tinytex)
library(tidyverse)
```

# 1. USARRESTS

## (a) Construct box plots for each of the four variables and comment on the univariate characteristics of each briefly.

```{r cache=TRUE}
data("USArrests")

# Construct box plots in a 2x2 grid
par(mfrow = c(2, 2), oma = c(0, 2, 0, 0))  # Set outer margin for the title
boxplot(USArrests$Murder, main = "Murder")
boxplot(USArrests$Assault, main = "Assault")
boxplot(USArrests$UrbanPop, main = "UrbanPop")
boxplot(USArrests$Rape, main = "Rape")

title("Box Plots for USArrests Variables", outer = TRUE, line = -1)

```

```{r cache=TRUE}
# Summary of univariate characteristics for each variable
summary(USArrests$Murder)
summary(USArrests$Assault)
summary(USArrests$UrbanPop)
summary(USArrests$Rape)
```

The "Murder" variable represents the rate of murder arrests per 100,000 individuals, providing insights into crime prevalence. Murder demonstrates a symmetric distribution centered around 7.250, with a range from 0.800 to 17.400 and no outliers. "Assault" indicates the rate of assault arrests per 100,000, shedding light on violent crime patterns. Assault exhibits a right-skewed distribution, centered at 159.0, with a range from 45.0 to 337.0 and no outliers. "UrbanPop" is the percentage of urban population, reflecting the urbanization level across states. UrbanPop displays near symmetry around 66.00, with a range of 32.00 to 91.00 and no outliers. Lastly, the "Rape" variable represents the rate of rape arrests per 100,000, offering an understanding of sexual assault occurrences. Rape is symmetric, centered at 20.10, with a range from 7.30 to 46.00, and there are some outliers.

## (b) Obtain the sample mean vector X; the sample covariance matrix S and the sample correlation matrix R: What can you say about the relationship between the four variables?

```{r cache=TRUE}
# Extracting the relevant columns
murder <- USArrests$Murder
assault <- USArrests$Assault
urbanPop <- USArrests$UrbanPop
rape <- USArrests$Rape

# Calculating the sample mean vector
mean_vector <- colMeans(USArrests)
cat("Sample Mean Vector (X):\n", mean_vector, "\n\n")

# Calculating the sample covariance matrix
cov_matrix <- cov(USArrests)
cat("Sample Covariance Matrix (S):\n")
print(cov_matrix)
cat("\n")

# Calculating the sample correlation matrix
cor_matrix <- cor(USArrests)
cat("Sample Correlation Matrix (R):\n")
print(cor_matrix)

```

-   Murder and Assault have a strong positive correlation (0.8018733).
-   Assault and Rape also show a strong positive correlation (0.6652412).
-   Murder and Rape exhibit a moderate positive correlation (0.5635788).
-   UrbanPop has weak correlations with the other variables.

In summary, the relationships between these variables suggest notable associations, especially between Murder, Assault, and Rape, as indicated by their respective.

## (c) Let... where S is the sample covariance matrix, be the statistical distance between an observed point x and a fixed point q. Find the state with the highest murder rate and compute the statistical distance of each observation from that state.

```{r cache=TRUE}
# Define the stat.dist function
stat.dist <- function(Y, p) {
  S.inv <- solve(var(Y))
  dist <- function(y) sqrt(t(y - p) %*% S.inv %*% (y - p))
  return(apply(Y, 1, dist))
}

# Convert USArrests to a matrix
X <- as.matrix(USArrests)

# Find the state with the highest murder rate
state_with_highest_murder_rate <- rownames(X)[which.max(X[, "Murder"])]

# Display the result
cat("State with the highest murder rate:", state_with_highest_murder_rate, "\n")

# Extract the data for the state with the highest murder rate
state_data <- X[state_with_highest_murder_rate, ]

# Calculate the statistical distance for each observation from the state with the highest murder rate
statistical_distances <- stat.dist(X, state_data)

# Create a data frame for better presentation
result_df <- data.frame(State = rownames(X), Distance = statistical_distances)

# Determine the maximum length of State names
max_state_length <- max(nchar(result_df$State))

# Display the results of Statistical distance

for (i in 1:nrow(result_df)) {
  cat(result_df$State[i], strrep(" ", max_state_length - nchar(result_df$State[i]) + 1), result_df$Distance[i], "\n")
}

```

## (d) Using these computed distances determine which six states are closest" to the state with the highest murder rate. Can you provide an explanation why these seven states are statistically similar in terms of their violent crime rates and percentage of urban population in 1973?

```{r cache=TRUE}
# Sort the data frame by statistical distances
sorted_result_df <- result_df[order(result_df$Distance), ]

# Select the top six closest states
closest_states <- sorted_result_df[2:7, ]

# Display the closest states
closest_states

```

```{r cache=TRUE}
# Subset the data for the seven states
selected_states <- c("Georgia", "Tennessee", "Louisiana", "Kentucky", "Texas", "Alabama", "Mississippi")
subset_data <- USArrests[selected_states, ]

# Display the subset of the data
subset_data

# Summary statistics for each variable
summary_stats <- apply(subset_data, 2, summary)
print(summary_stats)

# Arrange boxplots in a 2x2 grid
par(mfrow = c(2, 2))  # 
for (variable in colnames(subset_data)) {
  boxplot(subset_data[, variable], main = variable)
}
```

The seven states examined – Georgia, Tennessee, Louisiana, Kentucky, Texas, Alabama, and Mississippi – demonstrate statistical similarities in their crime rates and urban population percentages for the year 1973. Notably, the mean murder rate across these states is approximately 13.96, with individual states exhibiting rates ranging from 9.7 to 17.4. The mean assault rate is around 207.57, varying between 109 and 259 across states. Similarly, the mean rape rate hovers at approximately 22.14, with state-specific rates ranging from 16.3 to 26.9.

In terms of urban population, these states share a commonality, with an average percentage of around 59.86%. Individual states range from 44% to 80% in terms of urbanization, reflecting a relatively consistent urban population distribution.

This statistical coherence suggests a comparable landscape in both crime rates and urbanization levels across the examined states. While variations exist at the state level, the overall patterns in crime rates and urban population percentages align, possibly indicating shared socio-economic factors, cultural influences, or historical contexts. Further exploration and hypothesis testing would be necessary for a comprehensive understanding of the observed similarities.

```{r}
# Create a color vector for the points
point_colors <- ifelse(rownames(USArrests) %in% selected_states, "red", "blue")

# Create a matrix plot with customized points
pairs(USArrests, col = point_colors, pch = ifelse(rownames(USArrests) %in% selected_states, 16, 1))

```

# This is inline math: $y = mx + b$

This is display math:

$$ f(x) = \int_{a}^{b} x^2 \,dx $$

\`\`\`
